---
title: "Cold Water Swimming"
author: "Chris Gaskell"
date: "29/07/2021"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---

```{r setup, include=T, echo=FALSE, message=FALSE}
# Global options across code chunks
knitr::opts_chunk$set(echo = F, fig.width = 9, fig.height = 8, message = F, warning = F)

# Libraries
library(tidyverse)
library(scan)
library(SingleCaseES)
library(scales)
library(knitr)
library(kableExtra)
library(userfriendlyscience)
```

```{r libs, include=T, echo=FALSE, message=FALSE}

### Load required packages
require(userfriendlyscience)
safeRequire('ggplot2');
safeRequire('grid');
safeRequire('gridExtra');
safeRequire('viridis');
safeRequire('pander');
```

```{r data, include=F, results='hide'}
# Import the data as a csv file from your working directory. 
dat.long <- read.csv("CWS.csv") # Assuming that your data is stored in long format.

# Separate the nomothetic data
phq <- dat.long %>% filter(case == "PHQ9")

# Isolate the three idiographic measures
dat.long <- dat.long %>% filter(case == "SadHappy" | case == "AnxiousCalm" | case == "BoredExcited")

# rename the idiographic variables
dat.long <- dat.long %>% mutate(
    case = dplyr::recode(case,  "SadHappy" = "Happy",
                                "AnxiousCalm" = "Anxious",
                                "BoredExcited" = "Excited"))

dat.wide <- dat.long %>% spread(case, values) # Just incase we need to work in wide format.

# Make a Single Case Data Frame (special type of data format for working with data in the 'scan' package).
scdf <- readSC("CWS.csv")
scdf <- scdf[c(1:3)]
```

## About

This is an R markdown file to analyse the idiographic data collected from a SCED investigating the effectiveness of behavioural therapy (cold showers and cold water swimming) for major depressive disorder. The design for the study was an A-B-A-C-FU (or Baseline, Intervention 1, Withdrawal, Intervention 2, Follow-Up). The two intervention phases differed in that the former included cold shower, while the latter included both cold showers and cold water swimming. 

The three idiographic variables (i.e. dependent variables) represented self-report measures on a scale of 1-100. These included happiness (or conversely sad); excitement (or boredom); and finally anxiousness (or calmness). Positive change was indicated by an increase in scores for the sad-to-happy and board-to-excited measures, and by a decrease in scores for the calm-to-anxious measure.

An overview of the data structure is as follows: 
```{r inspect, include=T}
# Take a look at the data
#distinct(dat.long, case, phase, .keep_all = T)
```

```{r overview, include=F, results='hide'}
#summary(scdf)
```

# Hypotheses/pair-wise comparisons

Because the current study involved five phases (i.e. A-B-A-C-FU) there was subsequently as large number of potential pairwise comparison combinations which could be made. We selected four pairwise comparisons to fit with the hypotheses of the current study:
1. That introducing a basic behavioural treatment (i.e. cold showers) would lead to an improvement in depression measures (by comparing baseline with the first behavioural treatment).
2. That the treatment effects from the initial behavioural treatment would be not be sustained when the treatment is removed (by comparing withdrawal phase with original baseline phase).
3. That an additive treatment component of cold water swimming would lead to enhanced outcomes when compared with cold showers only (by comparing both treatment phases).
4. The the effects of treatment would be maintained by introduction of a relapse prevention plan (by comparing treatment 2 with the follow-up phase).

# Descriptive Statistics

Across the fives phases of the three idiographic measures there were no missing values.
The data-set was therefore complete with no imputation needed.
The means, standard deviations (SD), and trends are reported in the table.

```{r descriptives, include=T, warning=F, message=F}
desc <- describeSC(scdf)
desc <- desc$descriptives
desc <- desc %>% select(starts_with(c("m.","sd", "tren")))
desc <- round(desc, 2)

desc <- rename_with(desc, ~ tolower(gsub("m.", "mean_", .x, fixed = TRUE)))
desc <- rename_with(desc, ~ tolower(gsub("sd.", "sd_", .x, fixed = TRUE)))
desc <- rename_with(desc, ~ tolower(gsub("trend.", "trend_", .x, fixed = TRUE)))

desc <- desc %>% relocate((ends_with(c("a","b","a2","c","fu"))))
desc <- as.data.frame(desc)

desc <- t(desc)

knitr::kable(desc) %>% kable_styling("striped") %>%
  pack_rows("Baseline", 1, 3) %>%
  pack_rows("Cold Showers", 4, 6)%>%
  pack_rows("Withdrawal", 7, 9)%>%
  pack_rows("Cold Water", 10, 12)%>%
  pack_rows("Follow Up", 13, 15)

```




# Kendall's Tau

```{r tau, include=T, warning=F}
Tau <- list(
          A1vsB = tau_u(scdf, phases = c(1,2)),
          A1vsA2 = tau_u(scdf, phases = c(1,3)),
          BvsC = tau_u(scdf, phases = c(2,4)),
          CvsFU = tau_u(scdf, phases = c(4,5))
                         )
print(Tau, complete = T)
```

# Baseline analyses

The trends for the baseline period can be seen visually in the plots and statistically from the table.
When using Kendall's Tau to measure the trend of the baseline periods there was a significant trend for
happiness (t = `r round(Tau$A1vsB$table$SadHappy$Tau[2],2)`, p = `r format.pval(Tau$A1vsB$table$SadHappy$p[2], digits = 2, eps = .001)`) but not for
anxiousness (t = `r round(Tau$A1vsB$table$AnxiousCalm$Tau[2], 2)`, p = `r format.pval(Tau$A1vsB$table$AnxiousCalm$p[2], digits = 2, eps = .001)`) or
boredom (t = `r round(Tau$A1vsB$table$BoredExcited$Tau[2], 2)`, p = `r format.pval(Tau$A1vsB$table$BoredExcited$p[2], digits = 2, eps = .001)`).
As the trend for happiness showed a worsening, and the remaining variables did not show a significant trend, there was no need to apply statistical adjustment (i.e. for baseline trend) within the following analyses. 

The baseline and withdrawal phases are visually comparable. In terms of trends shown within the withdrawal period there was no significant trend shown for happiness (t = `r round(Tau$A1vsA2$table$SadHappy$Tau[3],2)`, p = `r format.pval(Tau$A1vsA2$table$SadHappy$p[3], digits = 2, eps = .001)`) or anxiousness (t = `r round(Tau$A1vsA2$table$AnxiousCalm$Tau[3], 2)`, p = `r format.pval(Tau$A1vsA2$table$AnxiousCalm$p[3], digits = 2, eps = .001)`)
however there was a significant trend for boredom (t = `r round(Tau$A1vsA2$table$BoredExcited$Tau[3], 2)`, p = `r format.pval(Tau$A1vsA2$table$BoredExcited$p[3], digits = 2, eps = .001)`). This indicates that during the withdrawal period there was some level of improvement across the phase.

In terms of level difference, Mann-whittney U tests were computed to determine is there was a statistically significant difference between the baseline and the withdrawal period. For happiness, there was a greater level of happiness shown in the withdrawal phase than the baseline phase. There was no significant difference for the Mann-whittney test for the boredom and anxiousness idiographics. Taken into account, there was small, limited evidence (statistically and visually) for differences between the baseline and withdrawal phases. 

# Non-overlap Statistics


```{r batchcalc, include=T, warning=F}
AvsB <- batch_calc_ES(dat.long, grouping = "case", condition = "phase", outcome = "values",
         ES = c("PND", "PEM", "NAP"), improvement = "increase",
         format = "wide", confidence = NULL, baseline_phase = "A", intervention_phase = "B")
AvsB <-  AvsB %>% as.data.frame(AvsB) %>% 
          mutate(across(c(2:4), percent, .01)) %>% 
          mutate(NAP_SE = round(NAP_SE, 2)) %>% 
          rename(PND = PND_Est, PEM = PEM_Est, NAP = NAP_Est)

AvsA2 = batch_calc_ES(dat.long, grouping = "case", condition = "phase", outcome = "values",
         ES = c("PND", "PEM", "NAP"), improvement = "increase",
         format = "wide", confidence = NULL, baseline_phase = "A", intervention_phase = "A2")
AvsA2 <-  AvsA2 %>% as.data.frame(AvsA2) %>% 
          mutate(across(c(2:4), percent, .01)) %>% 
          mutate(NAP_SE = round(NAP_SE, 2)) %>% 
          rename(PND = PND_Est, PEM = PEM_Est, NAP = NAP_Est)

BvsC = batch_calc_ES(dat.long, grouping = "case", condition = "phase", outcome = "values",
         ES = c("PND", "PEM", "NAP"), improvement = "increase",
         format = "wide", confidence = NULL, baseline_phase = "B", intervention_phase = "C")
BvsC <-  BvsC %>% as.data.frame(BvsC) %>% 
          mutate(across(c(2:4), percent, .01)) %>% 
          mutate(NAP_SE = round(NAP_SE, 2)) %>% 
          rename(PND = PND_Est, PEM = PEM_Est, NAP = NAP_Est)
 
CvsFU = batch_calc_ES(dat.long, grouping = "case", condition = "phase", outcome = "values",
        ES = c("PND", "PEM", "NAP"), improvement = "decrease",
        format = "wide", confidence = NULL, baseline_phase = "C", intervention_phase = "FU")
CvsFU <-  CvsFU %>% as.data.frame(CvsFU) %>% 
          mutate(across(c(2:4), percent, .01)) %>% 
          mutate(NAP_SE = round(NAP_SE, 2)) %>% 
          rename(PND = PND_Est, PEM = PEM_Est, NAP = NAP_Est)                 
 

nonover <- list(AvsB = AvsB, AvsA2 = AvsA2, BvsC = BvsC, CvsFU = CvsFU
                )

nonover

#label_percent()(nonover$AvsB)

#nonover$CvsFU %>% 
#  mutate(Test=Sepal.Length/45,Test=scales::percent(Test)) #
#
#scales::percent(nonover$AvsB[,-1])

```



```{r apa-plot, include=T}
plotSC(scdf,
       phase.names = c('Baseline', 'Cold-water', 'Withdrawal', 'Green spaces', 'Follow-up'),
       ylab= c("1-100"), # label for the y-axis
       xlab="Measurement days across the study", # label for the x-axis
       xinc = 10, # The increment number of the x axis
     lines=c('trend',lty='dotted',lwd=5,col='red'), # Lines to impose and styling.
     style=list(fill=TRUE,col.fill="white",pch=20), case.names = c("Happiness", "Anxiety", "Excited")) # Making it look pretty.
```


```{r plot, include=T, fig.height=6, fig.width=13}
Plot <- ggplot(dat.long, # replace 'dat.long' with your long format data frame.
          aes(x = mt, y = values)) + # replace x + y with your x and y variables.
          geom_line(aes(color = case # replace case with name for ideographic measure column.
                ), size = 1.3) 

Plot <- Plot +
          scale_color_manual(values = c("Darkgreen", "Darkblue", "red"))+
          ggtitle("Time series plot to demonstrate change across the three idiographic measures.")+
          jtools::theme_apa()+
          ylab("1 to 100 Rating Scale")+
          xlab("Measurement days across the study")

Plot <- Plot +
  geom_vline(xintercept = 28, color = "black", size=.25)+
  geom_vline(xintercept = 70, color = "black", size=.25)+
  geom_vline(xintercept = 98, color = "black", size=.25)+
  geom_vline(xintercept = 140, color = "black", size=.25) +
     annotate("text", x = .5*(0+25), y = 100, label = "Baseline", size=6) +
     annotate("text", x = .5*(25+70), y = 100, label = "Cold Showers", size=6) +
     annotate("text", x = .5*(70+98), y = 100, label = "Withdraw", size=6) +
     annotate("text", x = .5*(98+140), y = 100, label = "Cold Water Swim", size=6) +
     annotate("text", x = .5*(140+200), y = 100, label = "Follow-up", size=6)

Plot <- Plot +
  labs(caption = "Note. 100 is equal to (i) high happiness, (ii) high calmness, and (iii) high excitment.") +
  theme(
        axis.title=element_text(size=rel(60),face="bold"),
        plot.caption = element_text(hjust = 0, size = rel(1)),
        axis.text = element_text(size=rel(2)),
        plot.title = element_text(size=20),
        axis.title.x = element_text(size=20, face = "plain"),
        axis.title.y = element_text(size=20, face = "plain")
        )

Plot
```


# Regression Analyses
Both the piecewise and the generalised regression are analyses to determine change between phases. Bi-phasic comaprisons are possible but it does not seemt that more than 2 can be used meaningfully in a single model. The following regression analyses explores change following the introduction of the first intervention (i.e. A vs. B).


# Piecewise Regression

```{r pwr-data, fig.height=4}
# Import data
wide.dat <- dat.long %>%
          spread(case, values)

wide.dat <- wide.dat %>%
          filter(phase == "A" | phase == "B")

pwr_happy <- piecewiseRegr(wide.dat,
                timeVar='mt', yVar= 'Happy', phaseVar='phase',
                colors = list(
                         pre = "red",
                         post = "red",
                         diff = "Darkorange",
                         intervention = "black",
                         points = "black"),
                plotLabs = list(
                            title = "Happiness",
                                      caption = "Note. 100 is equal to (i) high happiness",
                                      subtitle = NULL,
                                      x = NULL,
                                      y = "Happiness (0-100)"),
                theme = theme(
                        panel.background = element_rect(fill = "white",
                        colour = "lightblue",
                        size = 0.5, linetype = "solid")),
                yBreaks = 3
              )

pwr_bored <- piecewiseRegr(wide.dat,
                timeVar='mt', yVar= 'Excited', phaseVar='phase',
                colors = list(
                         pre = "red",
                         post = "red",
                         diff = "Darkorange",
                         intervention = "black",
                         points = "black"),
                plotLabs = list(
                            title = "Excitement",
                                      caption = "Note. 100 is equal to (i) high excitement",
                                      subtitle = NULL,
                                      x = NULL,
                                      y = "Excitement (0-100)"),
                theme = theme(
                        panel.background = element_rect(fill = "white",
                        colour = "lightblue",
                        size = 0.5, linetype = "solid")),
                yBreaks = 3
              )

pwr_anxious <- piecewiseRegr(wide.dat,
                timeVar='mt', yVar= 'Anxious', phaseVar='phase',
                colors = list(
                         pre = "red",
                         post = "red",
                         diff = "Darkorange",
                         intervention = "black",
                         points = "black"),
                plotLabs = list(
                            title = "Anxiety",
                                      caption = "Note. 100 = high anxiety; regression = red line; level change = orange line",
                                      subtitle = NULL,
                                      x = NULL,
                                      y = "Anxiety (0-100)"),
                theme = theme(
                        panel.background = element_rect(fill = "white",
                        colour = "lightblue",
                        size = 0.5, linetype = "solid")),
                yBreaks = 3
              )



pwr.plots <- list(Happy = pwr_happy$output$plot,
                   Excited = pwr_bored$output$plot,
                   Anxious = pwr_anxious$output$plot)


```

```{r pwr-table, echo=FALSE, fig.height=4}

pwr_table <- data.frame(
    "measure" = c("Happy", "Excited", "Anxious"),
    "D" = c(pwr_happy$output$deviance, pwr_bored$output$deviance, pwr_anxious$output$deviance),
    "Rsq.null" = c(pwr_happy$output$Rsq.null, pwr_bored$output$Rsq.null, pwr_anxious$output$Rsq.null),
    "Rsq.mdl" = c(pwr_happy$output$Rsq.model, pwr_bored$output$Rsq.model, pwr_anxious$output$Rsq.model),
    "ESr" = c(pwr_happy$output$ES, pwr_bored$output$ES, pwr_anxious$output$ES),
    "Level change" = c(as.numeric(pwr_happy$output$coef[2]), as.numeric(pwr_bored$output$coef[2]), as.numeric(pwr_anxious$output$coef[2]),
    "Trend change" = c(as.numeric(pwr_happy$output$coef[4]), as.numeric(pwr_bored$output$coef[4]), as.numeric(pwr_anxious$output$coef[4]))
),
    "cohensD" = c(pwr_happy$intermediate$meanDiff$meanDiff.d, pwr_bored$intermediate$meanDiff$meanDiff.d, pwr_anxious$intermediate$meanDiff$meanDiff.d)
)

pwr_table <- pwr_table[1:3,]

pwr_table %>% mutate_at(2:7, round, 2)

```








# Generalised Logistic Regression


```{r glog-data, echo=FALSE, fig.height=4}
mycolors <- list("red", "red", "red", "red", "red")

glog.happy <- 
         genlog(wide.dat,
         timeVar='mt', yVar= 'Happy', phaseVar='phase',
         theme = theme(
                      panel.background = element_rect(fill = "white",
                      colour = "lightblue",
                      size = 0.5, linetype = "solid")),
         alphas = list(outsideRange = .0,
                       bounds = 0,
                       points = .0,
                       mid = 0),
         yBreaks = 3,
         colors = list(bottomBound = "black",
                       topBound = "black",
                       curve = "red",
                       mid = "Darkgreen",
                       intervention = "Darkblue",
                       points = "black",
                       outsideRange = "black"),
         plotLabs = list(title = "Happiness",
                        caption = "Note. 100 is equal to (i) high happiness",
                        subtitle = NULL,
                        x = NULL,
                        y = "Happiness (0-100)"))


glog.bored <- 
         genlog(wide.dat,
         timeVar='mt', yVar= 'Excited', phaseVar='phase',
         theme = theme(
                      panel.background = element_rect(fill = "white",
                      colour = "lightblue",
                      size = 0.5, linetype = "solid")),
         alphas = list(outsideRange = .0,
                       bounds = 0,
                       points = .0,
                       mid = 0),
         yBreaks = 3,
         colors = list(bottomBound = "black",
                       topBound = "black",
                       curve = "red",
                       mid = "Darkgreen",
                       intervention = "Darkblue",
                       points = "black",
                       outsideRange = "black"),
         plotLabs = list(title = "Excitement",
                        caption = "Note. 100 is equal to (i) high excitement",
                        subtitle = NULL,
                        x = NULL,
                        y = "Excitement (0-100)"))

glog.anxious <- 
         genlog(wide.dat,
         timeVar='mt', yVar= 'Anxious', phaseVar='phase',
         theme = theme(
                      panel.background = element_rect(fill = "white",
                      colour = "lightblue",
                      size = 0.5, linetype = "solid")),
         alphas = list(outsideRange = .0,
                       bounds = 0,
                       points = .0,
                       mid = 0),
         yBreaks = 3,
         colors = list(bottomBound = "black",
                       topBound = "black",
                       curve = "red",
                       mid = "Darkgreen",
                       intervention = "Darkblue",
                       points = "black",
                       outsideRange = "black"),
         plotLabs = list(title = "Anxiety",
                        caption = "Note. 100 = high anxiety; mid point = green line; intervention point = blue line",
                        subtitle = NULL,
                        x = NULL,
                        y = "Anxiety (0-100)"))


glog.plots <- list(Happy = glog.happy$output$plot,
                   Excited = glog.bored$output$plot,
                   Anxious = glog.anxious$output$plot)
```

```{r glog-table, echo=FALSE, fig.height=4}

glog.table <- data.frame(
    "measure" = c("Happy", "Excited", "Anxious"),
    "D" = c(glog.happy$output$deviance, glog.bored$output$deviance, glog.anxious$output$deviance),
    "Rsq.null" = c(glog.happy$output$Rsq, glog.bored$output$Rsq, glog.anxious$output$Rsq),
    "Rsq.mdl" = c(glog.happy$output$ESc, glog.bored$output$ESc, glog.anxious$output$ESc),
    "ESr" = c(glog.happy$output$ESr, glog.bored$output$ESr, glog.anxious$output$ESr),
    "growthRate" = c(glog.happy$output$growthRate, glog.bored$output$growthRate, glog.anxious$output$growthRate),
    "IP" = c(glog.happy$output$inflectionPoint, glog.bored$output$inflectionPoint, glog.anxious$output$inflectionPoint),
    "base" = c(glog.happy$output$base, glog.bored$output$base, glog.anxious$output$base),
    "top" = c(glog.happy$output$top, glog.bored$output$top, glog.anxious$output$top)

)

glog.table <- glog.table[1:3,]

glog.table %>% mutate_at(2:7, round, 2)

```





```{r plots, echo=FALSE, fig.height=13}


                   growplot = ggpubr::ggarrange(pwr.plots$Happy, glog.plots$Happy, pwr.plots$Excited, glog.plots$Excited, pwr.plots$Anxious, glog.plots$Anxious, nrow = 3, ncol = 2)
                   
                   growplots = ggpubr::annotate_figure(growplot,
                                                       top = ggpubr::text_grob("Piecewise regression plots                          Generalised logistic regression plots ", color = "Black", face = "bold", size = 16),
                                                       bottom = ggpubr::text_grob("Measurement days across the study", color = "Black", face = NULL, size = 14))
                   
                   growplots + ggpubr::rotate()

```











